---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive vocal patterns in schizophrenia? Report how many papers report quantitative estimates, comment on what percentage of the overall studies reviewed they represent (see PRISMA chart) your method to analyze them, the estimated effect size of the difference (mean effect size and standard error) and forest plots representing it. N.B. Only measures of pitch mean and pitch sd are required for the assignment (all the other measures have been removed from the dataset for the sake of simplicity). 

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

## Tips on the process to follow:

- Download the data on all published articles analyzing voice in schizophrenia and the prisma chart as reference of all articles found and reviewed
Data: https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0
Prisma chart: https://www.dropbox.com/s/vbjf6ff73dv9tru/PRISMA%202009%20flow%20diagram_PP.pptx?dl=0 
- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
    * Hint: PITCH_F0M and PITCH_F0SD group of variables are what you need
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

```{r}
# GUSTAV
# loading packages
pacman::p_load(lme4, lmerTest, readxl, tidyverse, metafor)
```

```{r}
# ANDERS
# loading meta analysis data
df <- read_excel('Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx')
```

```{r}
#SIGRID 
#### PREPPING COLUMNS IN DATAFRAME
df <- df %>% 
  mutate(
    SAMPLE_SIZE_HC = as.numeric(SAMPLE_SIZE_HC),
    SAMPLE_SIZE_SZ = as.numeric(SAMPLE_SIZE_SZ),
    PITCH_F0_HC_M = as.numeric(PITCH_F0_HC_M),
    PITCH_F0_SZ_M = as.numeric(PITCH_F0_SZ_M),
    PITCH_F0_HC_SD = as.numeric(PITCH_F0_HC_SD),
    PITCH_F0_SZ_SD = as.numeric(PITCH_F0_SZ_SD),
    TYPE_OF_TASK = as.factor(TYPE_OF_TASK)
  ) %>% 
  select(ArticleID, StudyID, Title, Authors, Year_publication, DIAGNOSIS, MALE_SZ, FEMALE_SZ, MALE_HC, FEMALE_HC, SAMPLE_SIZE_HC, SAMPLE_SIZE_SZ, TYPE_OF_TASK, PITCH_F0_HC_M, PITCH_F0_SZ_M, PITCH_F0_HC_SD, PITCH_F0_SZ_SD)
```


```{r}
############################# CALCULATING yi + vi ################################
# GUSTAV
# SMD = standardized mean difference, yi = observed effect size, vi = sampling variance
df <- escalc(
  measure = 'SMD', 
  n1i = SAMPLE_SIZE_HC, 
  n2i = SAMPLE_SIZE_SZ, 
  m1i = PITCH_F0_HC_M, 
  m2i = PITCH_F0_SZ_M, 
  sd1i = PITCH_F0_HC_SD, 
  sd2i = PITCH_F0_SZ_SD, data = df)

```

```{r}
############################## LMER MODELS ###############################
# ANDERS

m1_random <- lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data=df, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1_random)

m1_full <- lmer(yi ~ 1 + TYPE_OF_TASK + (1 | StudyID), weights = 1/vi, data=df, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1_full)
```

- Build a forest plot of the results (forest(model))
 
```{r} 
############################ RMA RANDOM EFFECTS MODEL ######################
# SIGRID

m2_random <-rma (yi, vi, data = df, slab = StudyID)
summary(m2_random)

# forest plot
forest(m2_random)

# baujat plot
baujat(m2_random)

```
 

```{r}
############################ RMA FIXED EFFECTS MODEL ##################################
#GUSTAV

# model with fixed and random effects
m2_full <- rma(yi, vi, mods = cbind(TYPE_OF_TASK), data = df, slab=StudyID)
summary(m2_full)

#forest plot
forest(m2_full)

#baujat plot
baujat(m2_full) 
```

```{r}
########################### FUNNEL PLOT AND NUMERIC TEST ##############################
# ANDERS

#funnel plots
funnel(m2_random)
funnel(m2_full)

#numeric tests
# regression model test
regtest(m2_random)
regtest(m2_full)

# correlation test - Kendall's Tau
ranktest(m2_random)
ranktest(m2_full)
```
 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

```{r}
# SIGRID
#loading dataset from assignment 3
data_3 <- read_csv("FINAL_DATA_A3.csv")
```
 

 
```{r}
############### ADDING STUDY FROM PORTFOLIO 3 TO THE DF #############################
# GUSTAV
#these dataframes contain values which are then added to the df
gender_diagnosis_data_3 <- data_3 %>% 
  group_by(Gender, Diagnosis) %>% 
  summarize(length(unique(New_id)))

sample_size_data_3 <- data_3 %>% 
  group_by(Diagnosis) %>% 
  summarize(length(unique(New_id)))

pitch_mean_data_3 <- data_3 %>% 
  group_by(Diagnosis) %>% 
  summarize(mean(mean)) 

pitch_sd_data_3 <- data_3 %>% 
  group_by(Diagnosis) %>% 
  summarize(sd(mean)) 

#adding a new row with the study from assignment 3
df <- as_tibble(df) %>% 
  add_row(ArticleID = 49,
          StudyID = 51,
          Title = 'Voice Patterns in Schizophrenia: A systematic Review and Bayesian Meta-Analysis',
          Authors = 'Parola Alberto, Simonsen Arndis, Bliksted Vibeke, Fusaroli Riccardo',
          DIAGNOSIS = "SZ",
          MALE_SZ = '73',
          FEMALE_SZ = '63',
          MALE_HC = '90',
          FEMALE_HC = '69',
          SAMPLE_SIZE_HC = 160,
          SAMPLE_SIZE_SZ = 136,
          TYPE_OF_TASK = "CONSTR",
          PITCH_F0_HC_M = 156.8,
          PITCH_F0_SZ_M = 167.0,
          PITCH_F0_HC_SD = 59.2,
          PITCH_F0_SZ_SD = 55.2
          )   

```

```{r}
############################# CALCULATING NEW VI + YI ########################
# ANDERS

df3 <- df %>% escalc(
  measure = 'SMD', 
  n1i = SAMPLE_SIZE_HC, 
  n2i = SAMPLE_SIZE_SZ, 
  m1i = PITCH_F0_HC_M, 
  m2i = PITCH_F0_SZ_M, 
  sd1i = PITCH_F0_HC_SD, 
  sd2i = PITCH_F0_SZ_SD, data = df)


df3$TYPE_OF_TASK <- as.factor(df3$TYPE_OF_TASK)

```

- Now look at the output of rma() and check tau and I2

```{r}
################ USING RMA (INCLUDING THE DATA FROM PORT 3) ###########################
# SIGRID

# first model with only random effects
m3_random <- rma(yi, vi, data = df3, slab = StudyID)
summary(m3_random)

# forest plot
forest(m3_random)

# baujat plot
baujat(m3_random) 

# Funnel plot
funnel(m3_random)


# second model with random and mixed effects
m3_full <- rma(yi, vi, mods = cbind(TYPE_OF_TASK), data = df3, slab=StudyID)
summary(m3_full)

# forest plot
forest(m3_full)

# baujat plot
baujat(m3_full) 
```
 

