---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive vocal patterns in schizophrenia? Report how many papers report quantitative estimates, comment on what percentage of the overall studies reviewed they represent (see PRISMA chart) your method to analyze them, the estimated effect size of the difference (mean effect size and standard error) and forest plots representing it. N.B. Only measures of pitch mean and pitch sd are required for the assignment (all the other measures have been removed from the dataset for the sake of simplicity). 

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

## Tips on the process to follow:

- Download the data on all published articles analyzing voice in schizophrenia and the prisma chart as reference of all articles found and reviewed
Data: https://www.dropbox.com/s/0l9ur0gaabr80a8/Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx?dl=0
Prisma chart: https://www.dropbox.com/s/vbjf6ff73dv9tru/PRISMA%202009%20flow%20diagram_PP.pptx?dl=0 

- Look through the dataset to find out which columns to use, and if there is any additional information written as comments (real world data is always messy!).
Hint: PITCH_F0M and PITCH_F0SD group of variables are what you need

```{r}
#loading packages
pacman::p_load(tidyverse, readxl, metafor, lmerTest)

```

```{r}
#Loading data
df <- read_excel("Matrix_MetaAnalysis_Diagnosis_updated290719.xlsx")

```


 
- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)
 
```{r}
########################################## preparing dataset and making yi and vi coloumns ######################################
df <- df %>% 
  mutate(
    SAMPLE_SIZE_HC = as.numeric(SAMPLE_SIZE_HC),
    SAMPLE_SIZE_SZ = as.numeric(SAMPLE_SIZE_SZ),
    PITCH_F0_HC_M = as.numeric(PITCH_F0_HC_M),
    PITCH_F0_SZ_M = as.numeric(PITCH_F0_SZ_M),
    PITCH_F0_HC_SD = as.numeric(PITCH_F0_HC_SD),
    PITCH_F0_SZ_SD = as.numeric(PITCH_F0_SZ_SD),
    TYPE_OF_TASK = as.factor(TYPE_OF_TASK)
  ) %>% 
  select(ArticleID, StudyID, Title, Authors, Year_publication, DIAGNOSIS, MALE_SZ, FEMALE_SZ, MALE_HC, FEMALE_HC, SAMPLE_SIZE_HC, SAMPLE_SIZE_SZ, TYPE_OF_TASK, PITCH_F0_HC_M, PITCH_F0_SZ_M, PITCH_F0_HC_SD, PITCH_F0_SZ_SD)
     
  

  
df_excl3 <- escalc(measure = "SMD", n1i = SAMPLE_SIZE_HC, n2i = SAMPLE_SIZE_SZ, m1i = PITCH_F0_HC_M, m2i = PITCH_F0_SZ_M,
             sd1i = PITCH_F0_HC_SD, sd2i = PITCH_F0_SZ_SD, data = df)
             
```




```{r}
################################################# using lmer models ##################################################

m1_only_random <- lmer(yi ~ 1 + (1 | StudyID), weights = 1/vi, data=df_excl3, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1_only_random)

m1_full <- lmer(yi ~ 1 + TYPE_OF_TASK+ (1 | StudyID), weights = 1/vi, data=df_excl3, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))
summary(m1_full)


```


```{r}
########################################## USING RMA ############################################

#first model with only random effects
m2_only_random <- rma(yi, vi, data = df_excl3, slab = StudyID)
summary(m2_only_random)
baujat(m2_only_random) #bajaut plot



#second model with random and mixed effects
m2_full <- rma(yi, vi, mods = cbind(TYPE_OF_TASK), data = df_excl3, slab=StudyID)
summary(m2_full)
baujat(m2_full) #bajaut plot



```
 
 
- Build a forest plot of the results (forest(model))
```{r}
############################ FOREST PLOTS ############################
forest(m2_only_random)
forest(m2_full)
``` 
 
 
 
 
Funnel Plot
```{r}
############################## FUNNEL PLOT AND NUMERIC TEST ##############################

#funnel plots
funnel(m2_only_random)
funnel(m2_full)


#numeric tests
regtest(m2_only_random)
regtest(m2_full)

ranktest(m2_only_random)
ranktest(m2_full)


```


 
- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?
```{r}
#loading dataset from assignment 3
data_3 <- read_csv("FINAL_DATA_A3.csv")
```
 

 
```{r}
################################### ADDING OUR OWN STUDY FROM PORT 3 TO THE DF ###########################################

#these dataframes inform values inputted in the df
#we manually insert values from these df's into the add_row argument in the next line of code
gender_diagnosis_data_3 <- data_3 %>% 
  group_by(Gender, Diagnosis) %>% 
  summarize(length(unique(New_id)))

sample_size_data_3 <- data_3 %>% 
  group_by(Diagnosis) %>% 
  summarize(length(unique(New_id)))

pitch_mean_data_3 <- data_3 %>% 
  group_by(Diagnosis) %>% 
  summarize(mean(mean)) 

pitch_sd_data_3 <- data_3 %>% 
  group_by(Diagnosis) %>% 
  summarize(sd(mean)) 




#adding a new row with our study from assignment 3
df <- as_tibble(df) %>% 
  add_row(ArticleID = 49,
          StudyID = 51,
          DIAGNOSIS = "SZ",
          MALE_SZ = 73,
          FEMALE_SZ = 63,
          MALE_HC = 90,
          FEMALE_HC = 69,
          SAMPLE_SIZE_HC = 160,
          SAMPLE_SIZE_SZ = 136,
          TYPE_OF_TASK = "CONSTR",
          PITCH_F0_HC_M = 156.8,
          PITCH_F0_SZ_M = 167.0,
          PITCH_F0_HC_SD = 59.2,
          PITCH_F0_SZ_SD = 55.2
          )   
          
          

```




```{r}
######################### RUNNING META ANALYSIS - (INCLUDING OUR DATA FROM PORT 3) ################################

df_incl3 <- escalc(measure = "SMD", n1i = SAMPLE_SIZE_HC, n2i = SAMPLE_SIZE_SZ, m1i = PITCH_F0_HC_M, m2i = PITCH_F0_SZ_M,
             sd1i = PITCH_F0_HC_SD, sd2i = PITCH_F0_SZ_SD, data = df)
```


- Now look at the output of rma() and check tau and I2
```{r}
########################################## USING RMA - (INCLUDING OUR DATA FROM PORT 3) ############################################

#first model with only random effects
m2_only_random <- rma(yi, vi, data = df_incl3, slab = StudyID)
summary(m2_only_random)
baujat(m2_only_random) #bajaut plot



#second model with random and mixed effects
m2_full <- rma(yi, vi, mods = cbind(TYPE_OF_TASK), data = df_incl3, slab=StudyID)
summary(m2_full)
baujat(m2_full) #bajaut plot



```
 
 
- Build a forest plot of the results (forest(model))
```{r}
############################ FOREST PLOTS INCLUDING PORT 3 - (INCLUDING OUR DATA FROM PORT 3)############################
forest(m2_only_random)
forest(m2_full)
``` 
```








