---
title: "Assignment 1 - Language Development in ASD - Power and simulations"
author: "[YOUR NAME]"
date: "[DATE]"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the third exciting part of the Language Development in ASD exercise

In this part of the assignment, we try to figure out how a new study should be planned (i.e. how many participants?) in order to have enough power to replicate the findings (ensuring our sample size is adequate, our alpha at 0.05 and our beta at 0.8):
1- if we trust the estimates of the current study. Report the power analysis and comment on what you can (or cannot) use its estimates for.
2- if we are skeptical of the current study. Report the power analysis and comment on what you can (or cannot) use its estimates for.
3- if we only have access to 30 participants. Identify the power for each relevant effect and discuss whether it's worth to run the study and why
The list above is also what you should discuss in your code-less report.


## Learning objectives

- Learn how to calculate statistical power
- Critically appraise how to apply frequentist statistical power

### Exercise 1

How much power does your study have (if your model estimates are quite right)?
- Load your dataset (both training and testing), fit your favorite model, assess power for your effects of interest (probably your interactions).
- Report the power analysis and comment on what you can (or cannot) use its estimates for.
- Test how many participants you would have to have to replicate the findings (assuming the findings are correct)

N.B. Remember that main effects are tricky once you have interactions in the model (same for 2-way interactions w 3-way interactions in the model). If you want to test the power of main effects, run a model excluding the interactions.
N.B. Check this paper: https://besjournals.onlinelibrary.wiley.com/doi/full/10.1111/2041-210X.12504
You will be using:
- powerSim() to calculate power
- powerCurve() to estimate the needed number of participants
- extend() to simulate more participants

```{r}
#loading packages
pacman::p_load(tidyverse, readr,dplyr,stringr,lmerTest,Metrics,caret, lme4, reshape2, simr)

```

```{r}
#loading function from assignment A2_P2
CleanUpData <- function(demo,LU,token){
  
  Speech <- merge(LU, token) %>% 
    rename(
      Subject = SUBJ, 
      Visit = VISIT) %>%
    mutate(
      Visit = as.numeric(str_extract(Visit, "\\d")),
      Subject = gsub("\\.","", Subject)
      ) %>%
    dplyr::select(
      Subject, Visit, MOT_MLU, CHI_MLU, types_MOT, types_CHI, tokens_MOT, tokens_CHI
    )
  
  demo <- demo %>%
    rename(
      Subject = Child.ID) %>% 
    dplyr::select(
      Subject, Visit, Ethnicity, Diagnosis, Gender, Age, ADOS, MullenRaw, ExpressiveLangRaw, Socialization
    ) %>%
    mutate(
      Subject = gsub("\\.","", Subject)
    )
    
  Data=merge(demo,Speech,all=T)
  
  Data1= Data %>% 
     subset(Visit=="1") %>% 
     dplyr::select(Subject, ADOS, ExpressiveLangRaw, MullenRaw, Socialization) %>%
     rename(Ados1 = ADOS, 
            verbalIQ1 = ExpressiveLangRaw, 
            nonVerbalIQ1 = MullenRaw,
            Socialization1 = Socialization) 
  
  Data=merge(Data, Data1, all=T) %>%
    mutate(
      Subject = as.numeric(as.factor(as.character(Subject))),
      Visit = as.numeric(as.character(Visit)),
      Gender = recode(Gender, 
         "1" = "M",
         "2" = "F"),
      Diagnosis = recode(Diagnosis,
         "A"  = "ASD",
         "B"  = "TD")
    )

  return(Data)
}
```




```{r}
#loading df and filtering out NA's from CHI_MLU
data <- read_csv("data_clean.csv") %>% 
  filter(
    is.na(CHI_MLU) == F, is.na(Age) == F) %>% 
  rename('Subject' = 'Child.ID')


#loading test data and cleaning it with function from A2_P2
data_test <- CleanUpData(read_csv("Demo_test.csv"),
                         read_csv("LU_test.csv"),
                         read_csv("token_test.csv"))
```

```{r}
#Fitting our favorite model. We have removed the growth curve because simr didnt like it.
#Also we changed "age" to "visit" because it is easier to interpret

#runnig model on full training data
m1 <- lme4::lmer(CHI_MLU ~ Diagnosis * Visit + (Visit|Subject), data) 
summary(m1)

#running model on test_data
m1_test <- lme4::lmer(CHI_MLU ~ Diagnosis * Visit + (Visit|Subject), data_test) 
summary(m1_test)

```


```{r}
#Accessing power for interaction
powerI <- powerSim(m1, simr::fixed('Diagnosis:Visit'), nsim = 50)
powerI
#power is 100% (93-100 confidence interval) with full dataset

#power analysis for test data pretending it is a small pilot study
powerI_test <- powerSim(m1_test, simr::fixed('Diagnosis:Visit'), nsim = 50)
powerI_test
#power is 30.00% (17.86, 44.61) with only 6 participants

```

```{r}
#How many participants are needed for for finding an effect?

#full dataset
powerCurveI <-  powerCurve(m1, simr::fixed('Diagnosis:Visit'), along = 'Subject', nsim = 50)
powerCurveI
plot(powerCurveI)
```


```{r}
################################## PILOT STUDY ######################################################

# "pilot" dataset from test data
powerCurveI_test <-  powerCurve(m1_test, simr::fixed('Diagnosis:Visit'), along = 'Subject', nsim = 50)
powerCurveI_test
plot(powerCurveI_test)

# here we will see if the powercurve for the pilotstudy will approximate/resemble the power curve of the full data, when we extend the number of participants to the same amount.

m1_test_extend <- extend(m1_test, along="Subject", n = length(unique(df$Subject)))

powerCurveI_test_extend <-  powerCurve(m1_test_extend, simr::fixed('Diagnosis:Visit'), along = 'Subject', nsim = 50)
powerCurveI_test_extend
plot(powerCurveI_test_extend)

# We see that this powerplot is very similar to the powerplot for the full dataset for model m1. Thus extending the number of participants from a pilot study succesfully appriximate the power of the real data.
```



### Exercise 2

How would you perform a more conservative power analysis?
By lowering the effect size/ estimate, we need more participants to get the sufficient amount of power of 80%. 

- Identify and justify a minimum effect size for each of your relevant effects
- take the model from exercise 1 and replace the effects with the minimum effect size that you'd accept.
```{r}
#We only have our interaction effect, which is Diagnosis:Visit. With the current model for the full dataset, the estimate for Diagnosis:Visit is 0.25. This entails that the average TD child has a MLU of 1.5 syllables per utterance higher than the average ASD at the 6th visit. 

#We think a difference at Visit 6 of 1 syllbles per utterance would be a sufficient effect size. Thus we set the effect size of Diagnosis:Visit to 1/(number_of_visits) = 0.167. 


fixef(m1)["DiagnosisTD:Visit"] <- 1/(length(unique(df$Visit))) 
summary(m1)


```

- assess the power curve by Child.ID, identifying an ideal number of participants to estimate each effect

```{r}
powerCurveI_cons <-  powerCurve(m1, simr::fixed('Diagnosis:Visit'), along = 'Subject', nsim = 50)
powerCurveI_cons
plot(powerCurveI_cons)

#with a more conservative estimate we need 50 participants to get a certain power of 80%. 
```


- if your power estimates do not reach an acceptable threshold simulate additional participants and repeat the previous analysis
no need to do dis.


- Report the power analysis and comment on what you can (or cannot) use its estimates for. 
see report.







### Exercise 3

Assume you have only the resources to collect 30 kids (15 with ASD and 15 TDs). Identify the power for each relevant effect and discuss whether it's worth to run the study and why

```{r}
# First we extend to 30 partipants. Remember that our m1 has an estimate of 0.167 for Diagnosis:Visit, as we have changed it. 

m1_30 <- extend(m1, along="Subject", n = 30)
zzz <- getData(m1_30)

#extending the power
powerI_30 <- powerSim(m1_30, simr::fixed("Diagnosis:Visit"), nsim = 50)
powerI_30

#plotting the power along subject
powerCurveI_30 <-  powerCurve(m1, simr::fixed('Diagnosis:Visit'), along = 'Subject', nsim = 50)
powerCurveI_30
plot(powerCurveI_30)

#we conclude that 30 participants is not enough to find an estimate of 0.176 for Diagnosis:Visit with a power of 80%.
```


```{r}
############################# Solution: extending visit ################################

#we dont have enough money for more participants. But we have enough money for extending visits.
#Now we invetsegate how many visits needed to get sufficient power with n=30. 

m1_10visit <- extend(m1_30, along="Visit", n = 10)


powerCurveI_10visit <-  powerCurve(m1_10visit, simr::fixed('Diagnosis:Visit'), along = 'Visit', nsim = 50)
powerCurveI_10visit
plot(powerCurveI_10visit)


#  We know from the previous parts of the assignment that the effect of MLU is a best described by a growth curve model (rather than a linear model). This means that the effect will stagnate. Our model uses the linear method, why extending visit will give an inaccurate estimate of power.



# extending visit is maybe less cost-efficient than extending participants. Maybe it is better to spend money on more participants than more visits, because visit stagnate. 

```






