---
title: "ANDERS_A2_P2_LangASD_predictions_instructions"
author: "Anders Hjulmand"
date: "9/16/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the second exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time against some new data. Does the model generalize well?
Then we will learn to do better by cross-validating models and systematically compare them.

The questions to be answered (in a separate document) are:
1- Discuss the differences in performance of your model in training and testing data
2- Which individual differences should be included in a model that maximizes your ability to explain/predict new data?
3- Predict a new kid's performance (let's call him Bernie) and discuss it against expected performance of the two groups

## Learning objectives

- Critically appraise the predictive framework (contrasted to the explanatory framework)
- Learn the basics of machine learning workflows: training/testing, cross-validation, feature selections

## Let's go

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/5pc05mh5jwvdfjk/demo_test.csv?dl=0
* Utterance Length data: https://www.dropbox.com/s/eegu8fea2entdqv/LU_test.csv?dl=0
* Word data: https://www.dropbox.com/s/cf4p84mzn2p1bev/token_test.csv?dl=0

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
- create the test dataset (apply the code from assignment 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: google the functions "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())


```{r, include = FALSE}
pacman::p_load(readr,dplyr,stringr,lmerTest,Metrics,caret, tidyverse)
```


```{r, include = FALSE}
## Clean up function, included to inspire you, tidyverse magic

CleanUpData <- function(Demo,LU,Word){
  
  Speech <- merge(LU, Word) %>% 
    rename(
      Child.ID = SUBJ, 
      Visit=VISIT) %>%
    mutate(
      Visit = as.numeric(str_extract(Visit, "\\d")),
      Child.ID = gsub("\\.","", Child.ID)
      ) %>%
    dplyr::select(
      Child.ID, Visit, MOT_MLU, CHI_MLU, types_MOT, types_CHI, tokens_MOT, tokens_CHI
    )
  
  Demo <- Demo %>%
    dplyr::select(
      Child.ID, Visit, Ethnicity, Diagnosis, Gender, Age, ADOS, MullenRaw, ExpressiveLangRaw, Socialization
    ) %>%
    mutate(
      Child.ID = gsub("\\.","", Child.ID)
    )
    
  Data=merge(Demo,Speech,all=T)
  
  Data1= Data %>% 
     subset(Visit=="1") %>% 
     dplyr::select(Child.ID, ADOS, ExpressiveLangRaw, MullenRaw, Socialization) %>%
     rename(Ados1 = ADOS, 
            verbalIQ1 = ExpressiveLangRaw, 
            nonVerbalIQ1 = MullenRaw,
            Socialization1 = Socialization) 
  
  Data=merge(Data, Data1, all=T) %>%
    mutate(
      Child.ID = as.numeric(as.factor(as.character(Child.ID))),
      Visit = as.numeric(as.character(Visit)),
      Gender = recode(Gender, 
         "1" = "M",
         "2" = "F"),
      Diagnosis = recode(Diagnosis,
         "A"  = "ASD",
         "B"  = "TD")
    )

  return(Data)
}
```


```{r, include = FALSE}
################################### MODEL RMSE FOR TRAINING DATA ##################################

# Load training Data
df_train <- read_csv("data_clean.csv")

#- recreate the models you chose last time (just write the code again and apply it to Train Data)
m1 <- lmer(CHI_MLU ~ Diagnosis * Age + (Age|Child.ID)+I(Age^2), df_train)
summary(m1)

#- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
df_train_fit <- df_train %>% 
  filter(is.na(CHI_MLU) == F, is.na(Age) == F) %>% 
  mutate(FITTED = fitted(m1)) 

rmse(df_train_fit$CHI_MLU, df_train_fit$FITTED)
# We get an rmse of 0,34
```


```{r, include = FALSE}
######################################## LOADING AND CLEANING TEST DATA <##################################################

#- create the test dataset (apply the code from assignment 1 or my function to clean up the 3 test datasets)
# Test data
DEMO <- read_csv("demo_test.csv")
LU <- read_csv("LU_test.csv")
WORD <- read_csv("token_test.csv")

#using the nice function
df_test <- CleanUpData(DEMO, LU, WORD)
```


```{r, include = FALSE}
################################### MODEL RMSE FOR TEST DATA ##################################

#- test the performance of the models on the test data (Tips: google the functions "predict()")
df_test$predicted <- predict(m1, df_test, allow.new.levels = T)

#removing the na
df_test_exclude_na <- df_test %>% 
  filter(is.na(CHI_MLU) == F)

rmse(df_test_exclude_na$CHI_MLU, df_test_exclude_na$predicted)
#we get an rmse of 1.36

#- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())
predict(m1, newdata = df_test, allow.new.levels = T, interval = "confidence")
```




[HERE GOES YOUR ANSWER]

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Use cross-validation to compare your model from last week with the basic model (Child MLU as a function of Time and Diagnosis, and don't forget the random effects!)
- (Tips): google the function "createFolds";  loop through each fold, train both models on the other folds and test them on the fold)


- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test dat



```{r}
################################### FUNCTION FOR FOLDING ##############################

foldingfun <- function(k, mod){
  
folds <- createFolds(unique(df_train_fit$Child.ID), k = k, list = TRUE, returnTrain = FALSE)

trainRMSE <- rep(NA, k)
testRMSE <- rep(NA, k)

i = 1
for (fold in folds){
  train <- subset(df_train_fit, !(Child.ID %in% fold))
  test <- subset(df_train_fit, (Child.ID %in% fold))
  model <- lmer(mod, train)
  train$prediction <- fitted(model)
  test$prediction <- predict(model, test, allow.new.levels = TRUE)
  trainRMSE[i] <- rmse(train$CHI_MLU, train$prediction)
  testRMSE[i] <- rmse(test$CHI_MLU, test$prediction)
  i = i + 1
}


results <- tibble(trainRMSE, testRMSE)

return(results)
}
```


```{r}
################################### FUNCTION FOR REPEATING FOLDING #################################

more_foldings <- function(k){

#make folds from 1 to k
k_list <- rep(1:k)

#making empty lists of NA with length k
train_rmse_value <- rep(NA, k)
test_rmse_value <- rep(NA, k)

#model used           
model1 <- CHI_MLU ~ Diagnosis * Age + (Age|Child.ID)+I(Age^2)  

#we want to start at index 1
i = 1 

for (value in k_list){
  #Sum all the training/test rmse-values for this particular fold
  train_rmse_value[i] <- sum(foldingfun(value, model1)[1])
  test_rmse_value[i] <- sum(foldingfun(value, model1)[2])
  
  #dividing the sum of rmse-values with the number of folds at the current value
  train_rmse_value[i] <- train_rmse_value[i]/value
  test_rmse_value[i] <- test_rmse_value[i]/value
  i = i + 1
  
}

#return a df with mean test and training rmse-values for each k-fold session
results <- tibble(train_rmse_value, test_rmse_value)

return(results)
}
```



```{r}
################################# FOLDS ARE RANDOM ##################################

#we make several instances of the same number of folds, to show the small variation rmse-values, because the folds are not always the
mean_rmse_for_24_folds_1 <- more_foldings(20)
mean_rmse_for_24_folds_2 <- more_foldings(20)
mean_rmse_for_24_folds_3 <- more_foldings(20)
mean_rmse_for_24_folds_4 <- more_foldings(20)
```


```{r}
#defining number of folds
num_fol <- seq.int(1:20)

ggplot() + 
  geom_point(aes(y = mean_rmse_for_24_folds_1$train_rmse_value, x = num_fol)) +
  geom_point(aes(y = mean_rmse_for_24_folds_2$train_rmse_value, x = num_fol, color = "red")) + 
  geom_point(aes(y = mean_rmse_for_24_folds_3$train_rmse_value, x = num_fol, color = "blue")) + 
  geom_point(aes(y = mean_rmse_for_24_folds_4$train_rmse_value, x = num_fol, color = "red4")) +
  geom_point(aes(y = mean_rmse_for_24_folds_1$test_rmse_value, x = num_fol)) +
  geom_point(aes(y = mean_rmse_for_24_folds_2$test_rmse_value, x = num_fol, color = "red")) + 
  geom_point(aes(y = mean_rmse_for_24_folds_3$test_rmse_value, x = num_fol, color = "blue")) + 
  geom_point(aes(y = mean_rmse_for_24_folds_4$test_rmse_value, x = num_fol, color = "red4")) + 
  labs(x="Number of Folds", y = "Mean RMSE for training (lower) and test (upper)")

#What seems to be important is both how many folds you do and thatyou do several k-folds (we have done 4) and find an average between those
```


```{r}
###################################################### MAKING ALL MODELS ##########################################
# Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results

#Here we run the folding-function on different models

#simple model
model2 = CHI_MLU ~ Diagnosis + Age + (Age|Child.ID)+I(Age^2)
rmse_simple <- foldingfun(10, model2)

#fully complex model from last assignment
model1 <- CHI_MLU ~ Diagnosis * Age + (Age|Child.ID)+I(Age^2) 
rmse_complex_full <- foldingfun(10, model1)

#fully complex model without growth curve
model1_1 <- CHI_MLU ~ Diagnosis * Age + (Age|Child.ID)
rmse_complex_no_growthcurve <- foldingfun(10, model1_1)

#fully complex model without age as random slope
model1_2 <- CHI_MLU ~ Diagnosis * Age + (1|Child.ID)+I(Age^2)
rmse_complex_no_ageRslope <- foldingfun(10, model1_2)



```

```{r}
############################## COMPARING MEAN OF MODELS #####################################

#defining function for getting mean of the rmse of the models

meanfunction <- function(data){
  
#transposing (reversing rows and coloumns) so we can use function rowMeans()
data <- t(data)   
data <- as_tibble(data)  

#get the mean of training and testRMSE
data <- mutate(data, mean = rowMeans(data)) %>% 
mutate(Type = c("Train", "Test"))
  
  
}

#using the function to calculate means of each dataframe  
rmse_simple <- meanfunction(rmse_simple)
rmse_complex_full <- meanfunction(rmse_complex_full)
rmse_complex_no_growthcurve <- meanfunction(rmse_complex_no_growthcurve)
rmse_complex_no_ageRslope <- meanfunction(rmse_complex_no_ageRslope)

#combining all the means into one dataframe
rmse_comparison <- data.frame(
  Type = rmse_simple[12],
  simple_model = rmse_simple[11],
  complex_model = rmse_complex_full[11],
  complex_model_no_growthcurve = rmse_complex_no_growthcurve[11],
  complex_model_no_ageRslope = rmse_complex_no_ageRslope[11]
)

rmse_comparison
```





```{r}
# Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data

rmse(df_test_exclude_na$CHI_MLU, df_test_exclude_na$predicted)
#we get an rmse of 1.35 for actual test data

rmse_comparison$complex_model
#rmse is a lot lower for test-fold in cross-validation


```

[HERE GOES YOUR ANSWER]

### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.


```{r}
########################## BERNIE IS BETTER THAN TD ###################################

# Extracting Bernie
df_bernie <- df_test %>% filter(Child.ID == 2)

df_train_fit$Visit <- as.factor(df_train_fit$Visit)

#Average CHI_MLU for TD at each visit
df_train_TD <-  df_train_fit %>% 
  filter(df_train_fit$Diagnosis == "TD") %>% 
  group_by(Visit) %>% 
  summarize(average_TD = mean(CHI_MLU))


df_bernie <- merge(df_bernie, df_train_TD)

  
  
ggplot(df_bernie, aes(x = Visit, y = CHI_MLU)) + 
  geom_point(color = "blue") + 
  geom_point(aes(y = average_TD)) + 
  geom_smooth(method = lm, aes(color = "Bernie"), formula = y ~ x + I(x^2)) + 
  geom_smooth(method = lm, aes(x = Visit, y = average_TD, color = "Average_TD"), formula = y ~ x + I(x^2)) + 
  scale_colour_manual(name = "legend", values = c("black", "blue"))
  
abs(df_bernie$CHI_MLU-df_bernie$average_TD)

```


- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```{r}
############################## BERNIE IS A LOT BETTER THAN HIS AUTISTIC FELLAS AT VISIT 6 #########################################

#df_bernie$berniepredict <- predict(m1, df_bernie, allow.new.levels = T) #same as before 

mean(df_bernie$CHI_MLU-df_bernie$predicted)

mean(df_bernie$CHI_MLU)


```


[HERE GOES YOUR ANSWER]

### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

```{r}
#we have two models
summary(m1)
summary(m2)

anova(m1,m2)
#M1 has lowest AIC and BIC

testRMSE
testRMSE_s
mean(testRMSE, na.rm = T)
mean(testRMSE_s, na.rm = T)

#rmse for m1 is largest, so the information criteria and rmse dont go together

```



### OPTIONAL: Exercise 5): Using Lasso for model selection
Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!








