---
title: "FINAL_A2_P2_LangASD_predictions_instructions"
author: "Anders Wheelman, Gustav Helmet, Sigrid Snapfield"
date: "9/17/2020"
output: html_document:
  keep_md: True
---

---
title: "Assignment 2 - Language Development in ASD - Making predictions"
author: "Riccardo Fusaroli"
date: "August 9, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Welcome to the second exciting part of the Language Development in ASD exercise

In this exercise we will delve more in depth with different practices of model comparison and model selection, by first evaluating your models from last time against some new data. Does the model generalize well?
Then we will learn to do better by cross-validating models and systematically compare them.

The questions to be answered (in a separate document) are:
1- Discuss the differences in performance of your model in training and testing data
2- Which individual differences should be included in a model that maximizes your ability to explain/predict new data?
3- Predict a new kid's performance (let's call him Bernie) and discuss it against expected performance of the two groups

## Learning objectives

- Critically appraise the predictive framework (contrasted to the explanatory framework)
- Learn the basics of machine learning workflows: training/testing, cross-validation, feature selections

## Let's go

N.B. There are several datasets for this exercise, so pay attention to which one you are using!

1. The (training) dataset from last time (the awesome one you produced :-) ).
2. The (test) datasets on which you can test the models from last time:
* Demographic and clinical data: https://www.dropbox.com/s/5pc05mh5jwvdfjk/demo_test.csv?dl=0
* Utterance Length data: https://www.dropbox.com/s/eegu8fea2entdqv/LU_test.csv?dl=0
* Word data: https://www.dropbox.com/s/cf4p84mzn2p1bev/token_test.csv?dl=0

### Exercise 1) Testing model performance

How did your models from last time perform? In this exercise you have to compare the results on the training data and on the test data. Report both of them. Compare them. Discuss why they are different.

- recreate the models you chose last time (just write the model code again and apply it to your training data (from the first assignment))
- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())
- create the test dataset (apply the code from assignment 1 to clean up the 3 test datasets)
- test the performance of the models on the test data (Tips: google the functions "predict()")
- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())

```{r packages and data}
pacman::p_load(readr,dplyr,stringr,lmerTest,Metrics,caret, lme4, reshape2)

```


```{r, include = FALSE}
# ANDERS
## Cleaning up test data

CleanUpData <- function(demo,LU,token){
  
  Speech <- merge(LU, token) %>% 
    rename(
      Subject = SUBJ, 
      Visit = VISIT) %>%
    mutate(
      Visit = as.numeric(str_extract(Visit, "\\d")),
      Subject = gsub("\\.","", Subject)
      ) %>%
    dplyr::select(
      Subject, Visit, MOT_MLU, CHI_MLU, types_MOT, types_CHI, tokens_MOT, tokens_CHI
    )
  
  demo <- demo %>%
    rename(
      Subject = Child.ID) %>% 
    dplyr::select(
      Subject, Visit, Ethnicity, Diagnosis, Gender, Age, ADOS, MullenRaw, ExpressiveLangRaw, Socialization
    ) %>%
    mutate(
      Subject = gsub("\\.","", Subject)
    )
    
  Data=merge(demo,Speech,all=T)
  
  Data1= Data %>% 
     subset(Visit=="1") %>% 
     dplyr::select(Subject, ADOS, ExpressiveLangRaw, MullenRaw, Socialization) %>%
     rename(Ados1 = ADOS, 
            verbalIQ1 = ExpressiveLangRaw, 
            nonVerbalIQ1 = MullenRaw,
            Socialization1 = Socialization) 
  
  Data=merge(Data, Data1, all=T) %>%
    mutate(
      Subject = as.numeric(as.factor(as.character(Subject))),
      Visit = as.numeric(as.character(Visit)),
      Gender = recode(Gender, 
         "1" = "M",
         "2" = "F"),
      Diagnosis = recode(Diagnosis,
         "A"  = "ASD",
         "B"  = "TD")
    )

  return(Data)
}

```


```{r, include = FALSE}
# SIGRID
##################################### Training data #####################################
# Load training Data
data_clean <- read_csv('data_clean.csv') 
data_clean <- data_clean %>% rename('Subject' = 'Child.ID')
data <- filter(data_clean, is.na(CHI_MLU) == F, is.na(Age) == F)

#- recreate the models you chose last time (just write the code again and apply it to Train Data)
m1 <- lmer(CHI_MLU ~ Diagnosis * Age + (Age|Subject) + I(Age^2), data_clean)
summary(m1)

#- calculate performance of the model on the training data: root mean square error is a good measure. (Tip: google the function rmse())

data <- data %>% 
  mutate(Fitted = fitted(m1))

rmse(data$CHI_MLU, data$Fitted)
```


```{r, include = FALSE}
# GUSTAV 
####################################### Test data #################################

#- create the test dataset (apply the code from assignment 1 or my function to clean up the 3 test datasets)
# Test data
data_test <- CleanUpData(read_csv("Demo_test.csv"),
                         read_csv("LU_test.csv"),
                         read_csv("token_test.csv"))

#- test the performance of the models on the test data (Tips: google the functions "predict()")

data_test <- data_test %>% 
  filter(is.na(CHI_MLU) == F, is.na(Age) ==F)


data_test <- data_test %>% 
  mutate(
    Predict = predict(m1, data_test, allow.new.levels = T))

rmse(data_test$CHI_MLU, data_test$Predict)

#- optional: predictions are never certain, can you identify the uncertainty of the predictions? (e.g. google predictinterval())

```

See report 

### Exercise 2) Model Selection via Cross-validation (N.B: ChildMLU!)

One way to reduce bad surprises when testing a model on new data is to train the model via cross-validation. 

In this exercise you have to use cross-validation to calculate the predictive error of your models and use this predictive error to select the best possible model.

- Use cross-validation to compare your model from last week with the basic model (Child MLU as a function of Time and Diagnosis, and don't forget the random effects!)
- (Tips): google the function "createFolds";  loop through each fold, train both models on the other folds and test them on the fold)


- Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results.

- Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?
- Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data


```{r}
# ANDERS 
#- Create the basic model of ChildMLU as a function of Time and Diagnosis (don't forget the random effects!).

m0 <- lmer(CHI_MLU ~ Diagnosis + Age + (Age|Subject), data)

```


```{r cross validation function}
# SIGRID

# Cross validation function
cross_valid <- function(data, k, mod){
  folds <- createFolds(unique(data$Subject), k = k, list = T, returnTrain = F)

trainRMSE <- rep(NA,k)
testRMSE <- rep(NA,k)

i = 1
for (fold in folds){
  train = subset(data, !(Subject %in% fold))
  test = subset(data, Subject %in% fold)
  model = lmer(mod, train)
  test$prediction <- predict(model, test, allow.new.levels = T)
  train$prediction <- fitted(model)
  trainRMSE[i] <- rmse(train$CHI_MLU, fitted(model))
  testRMSE[i] <- rmse(test$CHI_MLU, test$prediction)
  i = i+1
}
result = data.frame(trainRMSE, testRMSE)
  return (result)
}
```

```{r}
# GUSTAV 

#- Make a cross-validated version of the model. (Tips: google the function "createFolds";  loop through each fold, train a model on the other folds and test it on the fold)

m0_cv <- cross_valid(data, 6, m0)
m0_cv

#- Report the results and comment on them. 

mean(m0_cv$trainRMSE)
mean(m0_cv$testRMSE)

# The value of RMSE of the training data is 0.35, which is lower than the RMSE of the test data, which is 0.86.

```


```{r}
# ANDERS 

#############################3 MAKING ALL MODELS #################################
# Now try to find the best possible predictive model of ChildMLU, that is, the one that produces the best cross-validated results

#Here we run the folding-function on different models

#

m1 <-  CHI_MLU ~ Diagnosis + Age + (1|Subject)
rmse_m1 <- cross_valid(data, 10, m1)

m2 <-  CHI_MLU ~ Diagnosis + Age + (Age|Subject)
rmse_m2 <- cross_valid(data, 10, m2)

m3 <- CHI_MLU ~ Diagnosis + Age + (1|Subject)+ I(Age^2) 
rmse_m3 <- cross_valid(data, 10, m3)
  
m4 <- CHI_MLU ~ Diagnosis + Age + (Age|Subject)+ I(Age^2) 
rmse_m4 <- cross_valid(data, 10, m4)

m5 <- CHI_MLU ~ Diagnosis * Age + (1|Subject)
rmse_m5 <- cross_valid(data, 10, m5)

m6 <- CHI_MLU ~ Diagnosis * Age + (Age|Subject)
rmse_m6 <- cross_valid(data, 10, m6)

m7 <- CHI_MLU ~ Diagnosis * Age + (1|Subject)+I(Age^2)
rmse_m7 <- cross_valid(data, 10, m7)

m8 <- CHI_MLU ~ Diagnosis * Age + (Age|Subject)+I(Age^2) 
rmse_m8 <- cross_valid(data, 10, m8)

```

```{r}
# SIGRID

############################## COMPARING MEAN OF MODELS ##############################

#defining function for getting mean of the rmse of the models
meanfunction <- function(data){
  
#transposing (reversing rows and coloumns) so we can use function rowMeans()
data <- t(data)   
data <- as_tibble(data)  

#get the mean of training and testRMSE
data <- mutate(data, mean = rowMeans(data)) %>% 
mutate(Type = c("Train", "Test"))
  
  
}

#using the function to calculate means of each dataframe  
rmse_m1 <- meanfunction(rmse_m1)
rmse_m2 <- meanfunction(rmse_m2)
rmse_m3 <- meanfunction(rmse_m3)
rmse_m4 <- meanfunction(rmse_m4)
rmse_m5 <- meanfunction(rmse_m5)
rmse_m6 <- meanfunction(rmse_m6)
rmse_m7 <- meanfunction(rmse_m7)
rmse_m8 <- meanfunction(rmse_m8)


#combining all the means into one dataframe
rmse_comparison <- tibble(
  Type = rmse_m1[12],
  rmse_m1[11],
  rmse_m2[11],
  rmse_m3[11],
  rmse_m4[11],
  rmse_m5[11],
  rmse_m6[11],
  rmse_m7[11],
  rmse_m8[11],
)

t(rmse_comparison)

# model 8 has the lowest RMSE value for the test data.

```


```{r}
# GUSTAV 

# Bonus Question 1: What is the effect of changing the number of folds? Can you plot RMSE as a function of number of folds?

#number of folds
k = 20

#Making and empty dataframe for the output
cross_mean <- data.frame(
  k_fold = rep(1:k),
  TrainRMSE_mean = rep(NA,k), 
  TestRMSE_mean = rep(NA,k)
  )

#Making a loop 
for (i in 1:k){
  fold = cross_valid(data, i,m8)
  cross_mean[i,2] = mean(fold$trainRMSE)
  cross_mean[i,3] = mean(fold$testRMSE)
}

#Plotting the result
cross_mean %>% 
  melt(id.vars = "k_fold") %>% 
  ggplot()+
  aes(x = k_fold, y = value, colour = variable)+
  geom_point()+
  geom_smooth(method = lm) + labs(title = 'RMSE according to number of folds')
```



```{r}
# ANDERS 

# Bonus Question 2: compare the cross-validated predictive error against the actual predictive error on the test data

###Calculating the predictive error for the test data.###

#Creating the model
m1 <- lmer(CHI_MLU ~ Diagnosis*Age+(Age|Subject)+I(Age^2),data) 
#predicted values
predicted <- predict(m1, newdata = data_test, allow.new.levels = T) 

#The mean of the residuals between the predicted and actual data.
mean(abs(predicted-data_test$CHI_MLU))

###THE MEAN ERROR FOR THE ACTUAL TEST DATA  =  0.9893106###



###Calculating the predictive error for the cross-validated model###
k = 6
folds = createFolds(unique(data$Subject), k = k, list = T, returnTrain = F)

hep <- data.frame(trainRMSE = rep(NA, k),
                  testRMSE = rep(NA, k),
                  predictive_error = rep(NA,k)) #Adding a new collumn for mean residual

i = 1

for (fold in folds) { 
  train = subset(data, !(Subject %in% fold))
  test = subset(data, Subject %in% fold)
  model = lmer(CHI_MLU ~ Diagnosis*Age+(Age|Subject)+I(Age^2), train)
  test$prediction = predict(model, test, allow.new.levels = T)
  train$prediction = fitted(model)
  hep$trainRMSE[i] = rmse(train$CHI_MLU, fitted(model))
  hep$testRMSE[i] = rmse(test$CHI_MLU, test$prediction)
  hep$predictive_error[i] = mean(abs(test$prediction-data_test$CHI_MLU)) 
  i = i+1
}

#Mean of means : )
mean(hep$predictive_error)
###THE MEAN ERROR FOR THE CROSS VALIDATED MODEL ### # 0.7981
```

See report.

### Exercise 3) Assessing the single child

Let's get to business. This new kiddo - Bernie - has entered your clinic. This child has to be assessed according to his group's average and his expected development.

Bernie is one of the six kids in the test dataset, so make sure to extract that child alone for the following analysis.

You want to evaluate:

- how does the child fare in ChildMLU compared to the average TD child at each visit? Define the distance in terms of absolute difference between this Child and the average TD.


```{r}
# SIGRID 

########################## BERNIE IS BETTER THAN TD ###################################

#Getting the mean for each TD child pr visit
mean_td <- data %>% 
  filter(Diagnosis == "TD") %>% 
  group_by(Visit) %>% 
  summarise(mean_CHI_MLU = mean(CHI_MLU))

#Getting the mean for each ASD child pr visit for the plot
mean_asd <- data %>% 
  filter(Diagnosis == "ASD") %>% 
  group_by(Visit) %>% 
  summarise(mean_CHI_MLU = mean(CHI_MLU))

# Extracting Bernie
bernie <- data_test %>% 
  filter(Subject == 2) %>% 
  mutate(
    Absolute_dif = abs(mean_td$mean_CHI_MLU - CHI_MLU),
    mean_td = (mean_td$mean_CHI_MLU),
    mean_asd = (mean_asd$mean_CHI_MLU)) %>% 
  select(
    Subject, Diagnosis, Visit, CHI_MLU, mean_td, mean_asd, Age) 
  
#Plotting
bernie %>% 
  melt(id.vars = c("Subject","Visit","Age","Diagnosis")) %>% 
  ggplot()+
  aes(x = Visit, y = value, color = variable)+
  geom_point()+
  geom_smooth(method = lm, formula = y ~ x + I(x^2))

# PLOTTTING
bernie %>% 
  ggplot()+
  aes(x = Visit, y = Absolute_dif)+
  geom_point()+
  geom_smooth(method = lm, alpha = 0.1)


```

- how does the child fare compared to the model predictions at Visit 6? Is the child below or above expectations? (tip: use the predict() function on Bernie's data only and compare the prediction with the actual performance of the child)

```{r}
# GUSTAV 

# creating model 
m1 <- lmer(CHI_MLU ~ Diagnosis*Visit+(Visit|Subject)+I(Visit^2), data)

bernie <- bernie %>% 
  mutate(
    predicted = predict(m1, bernie, allow.new.levels = T)
  )

#according to the model, Bernie's predicted MLU at visit 6 is 1.5988

bernie %>% 
  filter(Visit == 6) %>% 
  select(Visit, CHI_MLU, predicted) %>% 
  mutate(
    dif = CHI_MLU - predicted
  )

# Bernie is better 

```


### OPTIONAL: Exercise 4) Model Selection via Information Criteria
Another way to reduce the bad surprises when testing a model on new data is to pay close attention to the relative information criteria between the models you are comparing. Let's learn how to do that!

Re-create a selection of possible models explaining ChildMLU (the ones you tested for exercise 2, but now trained on the full dataset and not cross-validated).

Then try to find the best possible predictive model of ChildMLU, that is, the one that produces the lowest information criterion.

- Bonus question for the optional exercise: are information criteria correlated with cross-validated RMSE? That is, if you take AIC for Model 1, Model 2 and Model 3, do they co-vary with their cross-validated RMSE?

### OPTIONAL: Exercise 5): Using Lasso for model selection
Welcome to the last secret exercise. If you have already solved the previous exercises, and still there's not enough for you, you can expand your expertise by learning about penalizations. Check out this tutorial: http://machinelearningmastery.com/penalized-regression-in-r/ and make sure to google what penalization is, with a focus on L1 and L2-norms. Then try them on your data!

