---
title: "ANDERS_A3_P2"
author: "Anders Hjulmand"
date: "10/21/2020"
output: html_document
---
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
For the most part yes, when there arr no language differences. We checked for that in A3_P1. 

- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
1) pitch variavility
2) pause duration (only chinese)

- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.



- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?


- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

```{r}
pacman::p_load(pROC, tidyverse, tidymodels)
```


```{r}
#this dataset contains study 1-6, danish and chinece. Excluding japanese
df <- read_csv("FINAL_DATA.csv")



```


We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve).
You need to think carefully as to how we should (or not) use study and subject ID.
```{r}
# best acoustic feature is pitch variability
df <- df %>%
    mutate(
        pitch_variability = range_high-range_low,
        scaled_pitch_variability = scale(pitch_variability),
        Diagnosis = as.factor(Diagnosis),
        New_id = as.factor(New_id),
        Language = as.factor(Language)
        )

#making chinese dataset with pauseduration
df_chinese <- df %>% 
  filter(Language == "Chinese") %>% 
  mutate(
    npause_1 = ifelse(npause == 0, 1, npause),
    pause_duration = (dur_s - phonationtime_s)/npause_1
  )


#building logistic regression
m1 <- glm(Diagnosis ~ pitch_variability, data = df, family = 'binomial')
summary(m1)

#chinese is also shit
m1_chinese <- glmer(Diagnosis ~ pause_duration + (1|New_id), 
               df_chinese, family = 'binomial')
summary(m1_chinese)

```

```{r}
##############FUNCTION FOR GETTING PROBABILITIES ########################
Logit2Prop <- function(logit){
  odds = exp(logit)
  prop = odds/(1+odds)
  return(prop)
}
```


```{r}
### making confusion matrix
df <- df %>% 
  mutate(
    pred = Logit2Prop(predict(m1, re.form=NA)),
    predDiagnosis = ifelse(pred >= 0.5, "Schizophrenia", "Control"),
    predDiagnosis = as.factor(predDiagnosis)
  )

confusionMatrix(df$predDiagnosis, df$Diagnosis, positive = "Schizophrenia")
```

```{r}
######### MAKING ROC CURVE ##################
rocCurve <- roc(response = df$Diagnosis, predictor = as.numeric(df$predDiagnosis))
auc(rocCurve)
ci(rocCurve)
plot(rocCurve, legacy.axes = T)

#this is how you compare models
```


Then cross-validate the logistic regression and re-calculate performance on the testing folds. 
N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 

```{r}
############ Function for cross-validating - calculating performance on each single test fold ##################
cross_valid_single <- function(data, k, mod){
df1 <- groupdata2::fold(data, k = k, cat_col = "Diagnosis", id_col = "New_id") %>% arrange(.folds)
  

test_AUC <- rep(NA, k)



for (i in seq.int(1,k)){
  #making folds and model based on Participant
  train = subset(df1, .folds =! i)
  test = subset(df1, .folds == i)
  model = glmer(mod, family = 'binomial', train)
  
  
  #confusion matrix
    test <- test %>% 
    mutate(
    pred = Logit2Prop(predict(model, test, allow.new.levels = T)),
    predDiagnosis = ifelse(pred >= 0.5, "Schizophrenia", "Control"),
    predDiagnosis = as.factor(predDiagnosis)
  )

 confu <- confusionMatrix(test$predDiagnosis, test$Diagnosis, positive = "Schizophrenia")
  
  
  #making ROC-curves for both train and test data
  rocCurve_test = roc(response = test$Diagnosis, predictor = as.numeric(test$predDiagnosis))
  
  #calculating AUC for both training and test
  test_AUC[i] <- auc(rocCurve_test)

  
  print(confu)
}

result = data.frame(test_AUC)
return(result)



}



```


```{r}
############ Function for cross-validating - calculate overall performance  ##################
# We had to fold by study instead of participant for some reason...
cross_valid_overall <- function(data, k, mod){
  
folds <- createFolds(unique(data$Study), k = k, list = T, returnTrain = F)
  
datalist <- list()
i = 1


for (fold in folds){
  #making folds and model based on Participant
  train = subset(data, !(Study %in% fold))
  test = subset(data, Study %in% fold)
  model = glmer(mod, family = 'binomial', train)
  
  #getting predicted probabilities for test
test <- test %>% 
  mutate(
    pred = Logit2Prop(predict(model, test, allow.new.levels = T)),
    predDiagnosis = ifelse(pred >= 0.5, "Schizophrenia", "Control"),
    predDiagnosis = as.factor(predDiagnosis),
    Diagnosis = as.factor(Diagnosis)
  )
  
  
  datalist[[i]] <- test
  
  
  i = i + 1
}

result <- dplyr::bind_rows(datalist) %>% 
  select(Diagnosis, predDiagnosis)


return(result)

}

```

```{r}
############ Function for cross-validating - calculate overall performance - with new folding techniques ##################

cross_valid_overall_new <- function(data, k, mod){
  
#making the better folds/partitioning
df1 <- groupdata2::fold(data, k = k, cat_col = "Diagnosis", id_col = "New_id") %>% arrange(.folds) 


datalist <- list()
#i = 1


for (i in seq.int(1,k)){
  #making folds and model based on Participant and balanced by Diagnosis
  train = subset(df1, .folds =! i)
  test = subset(df1, .folds == i)
  model = glmer(mod, family = 'binomial', train)
  
  #getting predicted probabilities for test
test <- test %>% 
  mutate(
    pred = Logit2Prop(predict(model, test, allow.new.levels = T)),
    predDiagnosis = ifelse(pred >= 0.5, "Schizophrenia", "Control"),
    predDiagnosis = as.factor(predDiagnosis),
    Diagnosis = as.factor(Diagnosis)
  )
  
  
  
  datalist[[i]] <- test
  
  
  #i = i + 1
}

result <- dplyr::bind_rows(datalist) %>% 
  select(Diagnosis, predDiagnosis)


return(result)

}

```





```{r}
##########################MODEL FOR CV #####################################

m1_cross_valid <- Diagnosis ~ scaled_pitch_variability + (1|New_id) + (1|Study)
```


```{r}
################################# PERFORMANCE OF SINGLE FOLDS ##############################
#getting auc for each single test fold
m1_cv_single <- cross_valid_single(df, 6, m1_cross_valid)
```



```{r}
######################### OVERALL PERFORMANCE OF FOLDS #################################
#running the cross_valid_overall function. The output of this function is a dataframe with actual values of diagnosis and predicted values of diagnosis for each test fold. 
m1_cv_overall <- cross_valid_overall(df,6, m1_cross_valid)


#making roc and auc and confusion matrix for the overall performance of test folds
confusionMatrix(m1_cv_overall$predDiagnosis, m1_cv_overall$Diagnosis, positive = "Schizophrenia")

rocCurve_cv_overall <- roc(response = m1_cv_overall$Diagnosis, predictor = as.numeric(m1_cv_overall$predDiagnosis))
plot(rocCurve_cv_overall)
auc(rocCurve_cv_overall)




```

```{r}
######################### OVERALL PERFORMANCE OF FOLDS -  new folding techniques #################################
#running the cross_valid_overall function. The output of this function is a dataframe with actual values of diagnosis and predicted values of diagnosis for each test fold. 
m1_cv_overall_new <- cross_valid_overall_new(df,6, m1_cross_valid)


#making roc and auc and confusion matrix for the overall performance of test folds
confusionMatrix(m1_cv_overall_new$predDiagnosis, m1_cv_overall_new$Diagnosis, positive = "Schizophrenia")

rocCurve_cv_overall_new <- roc(response = m1_cv_overall_new$Diagnosis, predictor = as.numeric(m1_cv_overall_new$predDiagnosis))
plot(rocCurve_cv_overall_new)
auc(rocCurve_cv_overall_new)





```


```{r}
################################################# TIDYMODELLING ##############################################
#This chunk is partioning the data, preprocessing the features, training the model, and then validating the model.

################################ PARTIONING - RSAMPLE ############################
#splitting into training and test
df_split <- initial_split(df, prop = 0.8)
df_split

#cheacking the training and test sets
df_split %>% 
  training() %>% 
  glimpse()

df_split %>% 
  testing() %>% 
  glimpse()


######################## PREPROCESSING FEATURES - RECIPES ####################
#extracting training data, making the recipe, define pre-processing steps and applying steps to training data
my_recipe <- training(df_split) %>% 
  recipe(Diagnosis ~ .) %>% 
  step_scale(all_numeric()) %>% 
  step_center(all_numeric()) %>% 
  prep()

my_recipe

#execute preprocessing on training and test data, selecting our predicting coloumn
my_training <- juice(my_recipe) %>% select(c(Diagnosis,pitch_variability))

my_testing <- my_recipe %>% 
  bake(testing(df_split)) %>% 
  select(c(Diagnosis,pitch_variability))

#note that juice() is doing just the same as bake(). It is just more neat. so juice(my_recipe) is doing the same
#as bake(my_recipe, new_data = my_training)


################### MODEL TRAINING - PARSNIP #########################
#making 4 classification-algorithms
log_fit <- 
  logistic_reg() %>% 
  set_mode("classification") %>% 
  set_engine("glm") %>% 
  fit(Diagnosis ~ pitch_variability, my_training)
summary(log_fit)

svm_fit <- 
  svm_rbf() %>% 
  set_mode("classification") %>% 
  set_engine("kernlab") %>% 
  fit(Diagnosis ~ pitch_variability, my_training)
summary(svm_fit)

RandomForest <- 
  rand_forest() %>%
  set_mode("classification") %>% 
  set_engine("randomForest") %>%
  fit(Diagnosis ~ pitch_variability, data = my_training) 

BoostedTree <- 
  boost_tree() %>%
  set_mode("classification") %>% 
  set_engine("xgboost") %>%
  fit(Diagnosis ~ pitch_variability, data = my_training) 


################## MODEL VALIDATION - YARDSTICK ####################
#this return a dataframe containing a coloumn for predicted values of each classification algorithm.
#for each classification algorithm we get both the predicted diagnosis and the probability of being schizophrenia
test_results <- 
  my_testing %>% 
  as_tibble() %>% 
  mutate(
    #LOG_FIT
    #getting predicted diagnosis
    log_class = predict(log_fit, new_data = my_testing) %>% 
    pull(.pred_class),
    #getting probability of schizophrenia
    log_prop = predict(log_fit, new_data = my_testing, type = "prob") %>% 
      pull(.pred_Schizophrenia),
    
    #SVM_FIT
    #getting predicted diagnosis
    svm_class = predict(svm_fit, new_data = my_testing) %>% 
    pull(.pred_class),
    #getting probability of schizophrenia
    svm_prop = predict(svm_fit, new_data = my_testing, type = "prob") %>% 
      pull(.pred_Schizophrenia),
    
    #RANDOM_FOREST
    #getting predicted diagnosis
    rf_class = predict(RandomForest, new_data = my_testing) %>% 
    pull(.pred_class),
    #getting probability of schizophrenia
    rf_prop = predict(RandomForest, new_data = my_testing, type = "prob") %>% 
      pull(.pred_Schizophrenia),
    
    #BOOSTED_TREE
    #getting predicted diagnosis
    bt_class = predict(BoostedTree, new_data = my_testing) %>% 
    pull(.pred_class),
    #getting probability of schizophrenia
    bt_prop = predict(BoostedTree, new_data = my_testing, type = "prob") %>% 
      pull(.pred_Schizophrenia)
  )

#getting the performance of the models and plotting roc curves
metrics(test_results, truth = Diagnosis, estimate = log_class) %>% knitr::kable()
metrics(test_results, truth = Diagnosis, estimate = svm_class) %>% knitr::kable()
metrics(test_results, truth = Diagnosis, estimate = rf_class) %>% knitr::kable()
metrics(test_results, truth = Diagnosis, estimate = bt_class) %>% knitr::kable()



test_results %>% 
  roc_curve(Diagnosis, log_prop) %>% 
  autoplot()

test_results %>% 
  roc_curve(Diagnosis, svm_prop) %>% 
  autoplot()

test_results %>% 
  roc_curve(Diagnosis, rf_prop) %>% 
  autoplot()

test_results %>% 
  roc_curve(Diagnosis, bt_prop) %>% 
  autoplot()




```



N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0. 
N.N.B. you need to decide whether to calculate performance on each single test fold or save all the prediction for test folds in one dataset, so to calculate overall performance.

N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation? Study doesnt work, because of unequal proportion of Diagnosis in the two groups. 


N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)




