---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

```{r}
#loading packages
pacman::p_load(tidyverse, caret, lme4, lmerTest, effsize, pROC, reshape2, groupdata2, cvms)
```


```{r}
#loading data only including study 1-6, excluding the Japanese study (study 7)
data <- read_csv('FINAL_DATA.csv')

#making significant variables to predict diagnosis
data <- data %>% mutate(
  pitch_variability = range_high-range_low,
  pitch_variability_scaled = scale(pitch_variability),
  pitch_variability_scaled = as.numeric(pitch_variability_scaled),
  Diagnosis = as.factor(Diagnosis),
  New_id = as.numeric(as.factor(New_id))
)
```


```{r}
#Defining the function for converting estimates into probabilities. 
logit2prob <- function(logit){
    odds = exp(logit)
    prob = odds/(1+odds)
    return(prob)
}
```

We first want to build a logistic regression to see whether you can diagnose schizophrenia from your best acoustic feature. Let's use the full dataset and calculate the different performance measures (accuracy, sensitivity, specificity, PPV, NPV, ROC curve). You need to think carefully as to how we should (or not) use study and subject ID.

```{r}
# making three different models
# logistic model with random intercept for subject - full model
m1 <- glmer(Diagnosis ~ data$pitch_variability_scaled + (1|New_id), data, family = 'binomial')
summary(m1)

# random effect model
m2 <- glm(Diagnosis ~ (1|New_id), data, family = 'binomial')
summary(m2)

#fixed effect model
m3 <- glm(Diagnosis ~ pitch_variability_scaled, data, family = 'binomial')
summary(m3)

```



```{r}
# model performance parameters 
data <- data %>% 
  mutate(
    #predicted values for m1
  pred_m1 = logit2prob(predict(m1)),
  pred_m1 = ifelse(pred_m1 >= 0.5, 'Schizophrenia', 'Control'),
  pred_m1 = as.factor(pred_m1),
  
  #predicted values for m2
  pred_m2 = logit2prob(predict(m2)),
  pred_m2 = ifelse(pred_m2 >= 0.5, 'Schizophrenia', 'Control'),
  pred_m2 = as.factor(pred_m2),
  
  #predicted values for m3
  pred_m3 = logit2prob(predict(m3)),
  pred_m3 = ifelse(pred_m3 >= 0.5, 'Schizophrenia', 'Control'),
  pred_m3 = as.factor(pred_m3)
)

#making confusion matrices for all three models
confusionMatrix(data = data$pred_m1, reference = data$Diagnosis, positive = 'Schizophrenia')
confusionMatrix(data = data$pred_m2, reference = data$Diagnosis, positive = 'Schizophrenia')
confusionMatrix(data = data$pred_m3, reference = data$Diagnosis, positive = 'Schizophrenia')


#ROC curve
#m1
roc_m1 <- roc(response = data$Diagnosis, predictor = as.numeric(data$pred_m1))
auc(roc_m1)
ci(roc_m1)
plot(roc_m1, axes.legacy = T)

#m2
roc_m2 <- roc(response = data$Diagnosis, predictor = as.numeric(data$pred_m2))
auc(roc_m2)
ci(roc_m2)
plot(roc_m2, axes.legacy = T)

#m3
roc_m3 <- roc(response = data$Diagnosis, predictor = as.numeric(data$pred_m3))
auc(roc_m3)
ci(roc_m3)
plot(roc_m3, axes.legacy = T)

```

Then cross-validate the logistic regression and re-calculate performance on the testing folds. N.B. The cross-validation functions you already have should be tweaked: you need to calculate these new performance measures. Alternatively, the groupdata2 and cvms package created by Ludvig are an easy solution. 


```{r}
################################### Cross validation function ########################################
cross_valid <- function(dataset, k, mod){

dataset <- groupdata2::fold(dataset, k = k, cat_col = "Diagnosis", id_col = "New_id") %>% arrange(.folds) 
  
train_AUC <- rep(NA, k) 
test_AUC <- rep(NA, k)
sensitivity <- rep(NA,k)
specificity <- rep(NA,k)
PPV <- rep(NA,k)
NPV <- rep(NA,k)
accuracy <- rep(NA,k)

i = 1
for (fold in seq.int(1,k)){
  train = subset(dataset, .folds != fold)
  test = subset(dataset, .folds == fold)
  model = glmer(mod, family = 'binomial', train)

  #getting predicted probabilities for both training and test
  train <- train %>% mutate(
    prediction = logit2prob(fitted(model)),
    prediction = ifelse(prediction >= 0.5, 'Schizophrenia', 'Control'),
    prediction = as.factor(prediction)
    )
  
  test <- test %>% mutate(
    prediction = logit2prob(predict(model, test, allow.new.levels = T)),
    prediction = ifelse(prediction >= 0.5, 'Schizophrenia', 'Control'),
    prediction = as.factor(prediction)
    )
  
  #making a confusion matrix
  cm = confusionMatrix(test$prediction, reference = test$Diagnosis, positive = 'Schizophrenia')
  
  #making ROC-curves for both train and test data
  rocCurve_train = roc(response = train$Diagnosis, predictor = as.numeric(train$prediction))
  rocCurve_test = roc(response = test$Diagnosis, predictor = as.numeric(test$prediction))
  
  #making variables
  accuracy[i] = cm$overall[[1]] 
  sensitivity[i] = cm$byClass[[1]]
  specificity[i] = cm$byClass[[2]]
  PPV[i] = cm$byClass[[3]]
  NPV[i] = cm$byClass[[4]]
  train_AUC[i] <- auc(rocCurve_train)
  test_AUC[i] <- auc(rocCurve_test)

  i = i+1
}
#result = data.frame(train$prediction)
result = data.frame(accuracy, sensitivity, specificity, PPV, NPV, train_AUC, test_AUC)
  return (result)
}
```


```{r}
###### TESTING STUFF - to be deleted

data$New_id <- as.factor(data$New_id)
df1 <- groupdata2::fold(data, k = 6, cat_col = "Diagnosis", id_col = "New_id") %>% arrange(.folds) 

test <- cross_valid(data, 6, m1_cv)


# logistic model with random intercept for subject - full model
m1_cv <- Diagnosis ~ pitch_variability_scaled + (1|New_id)

# random effect model
m2_cv <- Diagnosis ~ (1|New_id)

#fixed effect model
m3_cv <- Diagnosis ~ pitch_variability_scaled
```
N.B. the predict() function generates log odds (the full scale between minus and plus infinity). Log odds > 0 indicates a choice of 1, below a choice of 0.
N.N.B. you need to decide whether calculate performance on each single test fold or save all the prediction for test folds in one datase, so to calculate overall performance.
N.N.N.B. Now you have two levels of structure: subject and study. Should this impact your cross-validation?
N.N.N.N.B. A more advanced solution could rely on the tidymodels set of packages (warning: Time-consuming to learn as the documentation is sparse, but totally worth it)



```{r}



```

