---
title: "Assignment 3 - Part 2 - Diagnosing Schizophrenia from Voice"
author: "Riccardo Fusaroli"
date: "October 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 - Part 2 - Diagnosing schizophrenia from voice

In the previous part of the assignment you generated a bunch of "features", that is, of quantitative descriptors of voice in schizophrenia. We then looked at whether we could replicate results from the previous literature.
We now want to know whether we can automatically diagnose schizophrenia from voice only, that is, relying on the set of features you produced last time, we will try to produce an automated classifier.
Again, remember that the dataset containst 7 studies and 3 languages. Feel free to only include Danish (Study 1-4) if you feel that adds too much complexity.

Issues to be discussed your report:
- Should you run the analysis on all languages/studies at the same time? 
- Choose your best acoustic feature from part 1. How well can you diagnose schizophrenia just using it?
- Identify the best combination of acoustic features to diagnose schizophrenia using logistic regression.
- Discuss the "classification" process: which methods are you using? Which confounds should you be aware of? What are the strength and limitation of the analysis?
- Bonus question: Logistic regression is only one of many classification algorithms. Try using others and compare performance. Some examples: Discriminant Function, Random Forest, Support Vector Machine, etc. The package caret provides them. 
- Bonus Bonus question: It is possible combine the output of multiple  classification models to improve classification accuracy. For inspiration see,
https://machinelearningmastery.com/machine-learning-ensembles-with-r/
 The interested reader might also want to look up 'The BigChaos Solution to the Netflix Grand Prize'

## Learning objectives
- Learn the basics of classification in a machine learning framework
- Design, fit and report logistic regressions
- Apply feature selection techniques

### Let's start

```{r}
#Importing packages
pacman::p_load(tidyverse, lme4, caret, pROC, tidymodels, groupdata2)

```


```{r}
#Importing data
df <- read_csv("FINAL_DATA.csv") 

# Designing relevant predictors
df <- df %>%
    mutate(
        pitch_variability = range_high-range_low,
        proportion_spoken = phonationtime_s/dur_s,
        speech_rate = speechrate_nsyll_dur,
        npause_1 = ifelse(npause == 0, 1, npause),
        pause_duration = (dur_s - phonationtime_s)/npause_1,
        Diagnosis = as.factor(Diagnosis),
        New_id = as.factor(as.numeric(as.factor(New_id)))
        ) 

#Defining function to calculate probabilities:
logit2prob <- function(logit){
  odds = exp(logit)
  prob = odds/(1+odds)
  return(prob)
}
```


```{r}
######### SPLITTING DATA INTO TEST AND TRAINING SETS #########
# Making sure to get the same results each time
set.seed(332)

#The strata function makes sure that no inbalance in diagnosis will occur
df_split <- partition(df, p=0.8, cat_col = "Diagnosis", id_col = "New_id", list_out = F)
df_train <- subset(df_split, .partitions == 1)
df_test <- subset(df_split, .partitions == 2)
```

```{r}
########## PREPROCECCING OF THE DATA - MAKING A RECIPE ###########
#The recipe
df_rec <- recipe(Diagnosis ~ pitch_variability + proportion_spoken + speech_rate + pause_duration, data = df_train) %>% 
  step_scale(all_numeric()) %>% 
  step_center(all_numeric()) %>% 
  step_impute_linear(all_numeric())

#The preperation
df_prep <- prep(df_rec, retain= T)

#The juice
df_juice <- juice(df_prep)


```








